Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-03-13 src/tex/ms.tex
update

% Waveform families and slight motivation of why we take Phenom model
Currently, three families of GW models are commonly used. They are the
effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity
(NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models
\citep{husa2016frequency,khan2016frequency}. EOB models are constructed by
mapping two masses onto an effective body under an effective metric; NR
surrogate models construct waveforms using combinations of NR waveforms; Phenom
models are formulated using specific ansatz and inspiral approximations. While
EOB and NR surrogate models give better waveform approximants, Phenom waveforms
can be produced much faster, hence it is used mostly in data analysis tasks that
requires many waveform generations. This advantage scales up in data analysis
tasks such as matched filtering and parameter estimation, where many waveforms
are required in each run. This motivates us to improve upon the current
framework of Phenom models, thus can retain the advantage of fast waveform
generation while improving the model's accuracy.

@@

The rest of the paper is structured as follows: In
Sec.~\ref{subsec:waveform_model}, we explain how IMRPhenomD is constructed and
have a detailed explanation of how waveform coefficients are introduced to the
model. We discuss the mismatch in Sec.~\ref{subsec:mismatch}, where we explained
how we calculate it in conjunction with NR waveforms. Then, we describe the
optimization scheme used for recalibration in Sec.~\ref{subsec:optimization}. In
Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
of optimized waveforms with original waveforms. We also show how the
optimization result differs with waveforms of different intrinsic parameters. In
Sec.~\ref{sec:discussion}, we address the difference between our calibrating
procedure with \citep{khan2016frequency}. We also explain how reduced spin
parameterization affects the accuracy of the model.

@@

IMRPhenomD is a frequency-domain phenomenological model for modeling
spin-aligned BBH coalescence events. It was developed as an overhaul of previous
Phenom models, such as  IMRPhenomB/C
\citep{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW
data analyses. 
% Using {\ripple} \citep{ripple} and {\jax} \citep{jax2018github}, we are able to easily recalibrate the model. 
IMRPhenomD is constructed in separate segments, i.e. inspiral, intermediate, and merger-ringdown. 
\begin{align}
	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta)
\end{align}
In each segment, the amplitude and phase are made using simple functions of
frequency such as polynomials or lorentzians. These simple analytic functions
consists of parameters $\Lambda^i$, which are defined as follows.
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)
\end{align}
$\Lambda^i$ depends on waveform coefficients $\lambda$ and intrinsic parameters
including symmetric mass ratio $\eta$ and post-newtonian spin parameter
$\chi_{\mathrm{PN}}$, 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2).
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the mass and spin of larger and smaller
black hole respectively. Finally, the individual segments are connected in a way
that the final waveform is continuous in its first derivative.

>>>>

% Waveform families and slight motivation of why we take Phenom model
Currently, three families of GW models are commonly used. They are the
effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity
(NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models
\citep{husa2016frequency,khan2016frequency} \kw{Add more citations related to
the newest models}. EOB models are constructed by mapping two masses onto an
effective body under an effective metric; NR surrogate models construct
waveforms using combinations of NR waveforms; Phenom models are formulated using
specific ansatz and inspiral approximations. While EOB and NR surrogate models
give better waveform approximants, Phenom waveforms can be produced much faster,
hence it is used mostly in data analysis tasks that requires many waveform
generations. This advantage scales up in data analysis tasks such as matched
filtering and parameter estimation, where many waveforms are required in each
run. This motivates us to improve upon the current framework of Phenom models,
thus can retain the advantage of fast waveform generation while improving the
model's accuracy.

@@

The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
review the parameterization of the IMRPhenomD model and the mismatch function
that is used as an objective function for the calibration, followed by outlining the
specific optimization scheme used for recalibration. In
Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
of optimized waveforms with original waveforms. We also show how the
optimization result differs with waveforms of different intrinsic parameters. In
Sec.~\ref{sec:discussion}, we address the difference between our calibrating
procedure with \citep{khan2016frequency}. We also explain how reduced spin
parameterization affects the accuracy of the model.

@@

In order to recalibrate the model, we have to understand what parameters the model has.
Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
For interested readers, please refer to \cite{khan2016frequency} for more details on construction of the model.

The IMRPhenomD model is constructed by combining three individually fitted parts
into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
\kw{Check if the following equation is correct. I think it is off by some smoothing factors.}
\begin{align}
	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta).
\end{align}

Instead of fitting the strain, which is a highly oscillatory function that is
difficult to fit, the amplitude and phase are fitted since they are smoother
functions. In each part, the amplitude and phase are made using simple functions
of frequency such as polynomials or lorentzians. These simple analytic functions
consists of parameters $\Lambda^i$, which are defined as follows. \kw{Can you
change equation one such that it consists of $\Lambda$? Also be explicit about
the form of the function used for the fitting. Like write down the polynomials
and lorentizans for the specific parts.}
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2),
\end{align}
where $\lambda$ are fitting coefficients obtained during calibration, $\eta$ is
the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
parameter, which is defined as 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2).
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
respectively. Finally, the individual segments are connected in a way that the
final waveform is continuous in its first derivative \kw{expand a bit}.
